---
title: "COMPSCIX 415.2 Homework 1"
author: "Katherine Leu"
date: "June 10, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

My Github repository for my assignments can be found at this URL: [https://github.com/leukat/compscix-415-2-assignments]


## Looking at the Data
```{r load_packages, warning=FALSE, message=FALSE}
library(mdsr)
library(tidyverse)
```

```{r}
data("WorldCities")
```

### Question 1: How many observations and variables are there in this data set? What are some of the variable names?

There are 23,018 observations and 10 variables in the WorldCities dataset.  Variable names include "latitude," "longitude," "population," and "region."
```{r}
glimpse(WorldCities)
```

## Extracting Parts of the Data
#### Example:
```{r}
WorldCities <- head(WorldCities, 200) # 200 rows
country_col <- WorldCities$country
unique(country_col)
```

### Question 2: There is a column called region in the data set. Can you extract this and show only the unique values?
```{r}
region_col <- WorldCities$region
```
Unique values of region:
```{r}
unique(region_col)
```

## Extracting Data the Tidy Way
### Question 3: Can you extract and show the unique entries from the country column in WorldCities using one line of code and two %>% operators?
```{r}
WorldCities %>% select(country) %>% unique()
```
## Visualizing the Data
#### Example for visualizing city populations:
```{r}
WorldCities %>% ggplot(aes(x = population)) + 
  geom_histogram()
```

### Question 4: Make one more histogram of the population and add a new x-label, a new y-label, a new title (use ggtitle('my title')), and change the theme of the plot using theme_bw().
```{r}
WorldCities %>% ggplot(aes(x = population)) + 
  geom_histogram() +
  xlab('City Population') +
  ylab('Number of Cities') +
  ggtitle('Distribution of City Populations') +
  theme_bw()
```

