---
title: "COMPSCIX 415.2 Homework 1"
author: "Katherine Leu"
date: "June 10, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

My Github repository for my assignments can be found at this URL: [https://github.com/leukat/compscix-415-2-assignments]

```{r load_packages, warning=FALSE, message=FALSE}
library(mdsr)
library(tidyverse)
```

```{r}
data("WorldCities")
```
### Question 1: How many observations and variables are there in this data set? What are some of the variable names?

There are 23,018 observations and 10 variables in the WorldCities dataset.  Variable names include "latitude," "longitude," "population," and "region."
```{r}
glimpse(WorldCities)
```

### Question 2: There is a column called region in the data set. Can you extract this and show only the unique values?

####Example of extracting part of the data:
```{r}
WorldCities <- head(WorldCities, 200) # 200 rows
country_col <- WorldCities$country
unique(country_col)
```
####Extracting part of the data for region:
```{r}
region_col <- WorldCities$region
```
Unique values of region:
```{r}
unique(region_col)
```
